---
title: "Pteropod Puget Sound time series"
author: ""
output: 
  html_document
self_contained: yes
runtime: shiny
---

```{r, message = F, warning = F, echo = F}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, dev.args = list(bg = 'transparent'), eval = T, echo = F)

library(tidyverse)
library(shiny)
library(gridExtra)
library(glue)
library(lubridate)
library(patchwork)
library(shinyWidgets)

source('R/funcs.R')

data(chmdatsum)
data(chmdatraw)
data(biodat)

# pteropod birthday
strt <- '2008-06-01' %>% 
  as.Date

# make cohortyr an ordered factor
biodat <- biodat %>% 
  mutate(cohortyr = factor(cohortyr, ordered = T))
chmdatsum <- chmdatsum %>% 
  mutate(cohortyr = factor(cohortyr, ordered = T))
```

```{r reactives}
# cumulative stress data
strdat <- reactive({

  # inputs
  val <- input$val
  sta <- input$sta
  thrsh <- input$thrsh

  out <- chmdatsum %>% 
    filter(var %in% 'ara') %>% 
    full_join(biodat, by = c('station', 'date', 'yr', 'cohortyr', 'mo')) %>% 
    select(-lon, -lat, -var, -abu, -avesz, -stdsz, -yr) %>% 
    gather('ara', 'chmval', ave:dlt) %>% 
    filter(ara %in% val) %>% 
    filter(station %in% sta) %>% 
    mutate(
      aracat = ifelse(chmval < thrsh, 1, 0),
      aradff = thrsh - chmval
      ) %>% 
    unite('costa', cohortyr, station, remove = F) %>% 
    group_by(costa) %>% 
    mutate(
      strsdis = cumsum(aracat),
      strscnt = cumsum(aradff)
    ) 

  return(out)
  
})

# wq chem labels
chmlab <- reactive({
  
  # input
  chm <- input$chm
  val <- input$val
  
  if(chm == 'araund')
    return(chm)
  
  out <- paste(val, chm)
  
  return(out)
  
})
```

```{r}
column(12, 
  column(4, 
    selectInput('chm', 'Select chemistry:', choices = c('ara', 'araund', 'co3', 'ph', 'pco2', 'temp', 'oxy'))
  ),
  
  column(4, 
    selectInput('val', 'Select chemistry value:', choices = c('min', 'ave', 'max', 'std', 'rng', 'dlt'))
  ),
  
  column(4, 
    selectInput('dis', 'Select dissolution:', choices = c('typ1', 'typ2', 'typ3'), selected = 'typ3')
  )
)
column(12, 
  column(4, 
  
    pickerInput(inputId = "sta", label = 'Select stations:', choices = unique(chmdatsum$station),
      options = list(`actions-box` = TRUE, size = 20), selected = unique(chmdatsum$station), multiple = TRUE)    
              
  )       
)
```

Year values shown correspond to "cohort-years" that begin in June with spring spawning, i.e., the April "cohort-year" is one year prior to the calendar year.  

```{r}
# data for first set of plots in first three tabs
disdat <- reactive({
  
  # inputs
  chm <- input$chm
  dis <- input$dis
  val <- input$val
  sta <- input$sta
  
  dat <- chmdatsum %>% 
    filter(var %in% chm) %>% 
    full_join(biodat, by = c('station', 'date', 'yr', 'cohortyr', 'mo')) %>% 
    select(-lon, -lat, -var, -abu, -avesz, -stdsz) %>% 
    gather('dissvar', 'dissval', typ1:typ3) %>% 
    gather('chemvar', 'chemval', ave:dlt) %>% 
    filter(dissvar %in% dis) %>% 
    filter(chemvar %in% val) %>% 
    filter(station %in% sta)  
  
  return(dat)
  
})

# regression mods for first set of plots in first three tabs
dismod <- reactive({
  
  # input
  disdat <- disdat()
  
  # complete model
  all <- lm(dissval ~ chemval, data = disdat)
  
  # by cohort years
  bycohortyr <- disdat %>% 
    group_by(cohortyr) %>% 
    nest %>% 
    mutate(
      mod = purrr::map(data, function(x){ try({lm(dissval ~ chemval, data = x)})})
    )
  bycohortyrout <- pull(bycohortyr, mod)
  names(bycohortyrout) <- pull(bycohortyr, cohortyr)

  # by seasons
  byseas <- disdat %>% 
    group_by(mo) %>% 
    nest %>% 
    mutate(
      mod = purrr::map(data, function(x){ try({lm(dissval ~ chemval, data = x)})})
    )
  byseasout <- pull(byseas, mod)
  names(byseasout) <- pull(byseas, mo)

  # combine all    
  out <- c(
    all = list(all), 
    bycohortyrout,
    byseasout
    )
  
  return(out)
  
})
```

## {.tabset}

### All data

This plot shows the selected pteropod response measure against the selected environmental measure observed at the time of sampling.

```{r}
renderPlot({
  
  # inputs
  disdat <- disdat()
  dis <- input$dis
  chmlab <- chmlab()

  p <- ggplot(disdat, aes(x = chemval, y = dissval)) +
    # geom_line(aes(group = station), colour = 'grey') +
    stat_smooth(method = 'lm', se = F, colour = 'red') +
    geom_label(aes(fill = mo, group = station, label = station), colour = 'grey', size = 3) +
    ylab(dis) +
    xlab(chmlab) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      legend.title = element_blank(), 
      legend.position = 'top'
    )

  return(p)
  
}, height = 400, width = 400)

renderPrint({lapply(dismod()['all'], summary)})
```

### Grouped by year

The first set of plots show the selected pteropod response measure against the selected environmental measure observed at the time of sampling.  Plots are facetted by cohort-year with points in each facet colored by month. 

```{r}
renderPlot({
  
  # inputs
  disdat <- disdat()
  dis <- input$dis
  chmlab <- chmlab()

  p <- ggplot(disdat, aes(x = chemval, y = dissval)) +
    # geom_line(aes(group = station), colour = 'grey') +
    stat_smooth(method = 'lm', se = F, colour = 'red') +
    geom_label(aes(fill = mo, group = station, label = station), colour = 'grey', size = 3) +
    facet_grid(~ cohortyr) + #, scales = 'free') +
    ylab(dis) +
    xlab(chmlab) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      legend.title = element_blank(), 
      legend.position = 'top'
    )

  return(p)
  
}, height = 300, width = 850)

renderPrint({lapply(dismod()[c('2014', '2015', '2016')], summary)})
```

The second set of plots show the selected environmental measure against temperature with points sized by the pteropod response measure. The x-axis is alway temperature and can be selected to show minimum, average, and maximum. Plots are facetted by cohort-year with points in each facet colored by month. 
```{r}
column(12, 
  column(4, 
    selectInput('tmp', 'Select x-axis temperature value:', choices = c('min', 'ave', 'max', 'std', 'rng', 'dlt'))
  )
)

renderPlot({
  
  validate(
    need(input$chm != 'temp', 'temperature on both plot axes not possible')
  )
  
  # inputs
  chm <- input$chm
  dis <- input$dis
  val <- input$val
  tmp <- input$tmp
  sta <- input$sta
  chmlab <- chmlab()

  dat <- chmdatsum %>% 
    filter(var %in% c(chm, 'temp')) %>% 
    full_join(biodat, by = c('station', 'date', 'yr', 'cohortyr', 'mo')) %>% 
    select(-lon, -lat, -abu, -avesz, -stdsz) %>% 
    gather('dissvar', 'dissval', typ1:typ3) %>% 
    gather('chemvar', 'chemval', ave:dlt) %>% 
    filter(dissvar %in% dis) %>% 
    filter(var %in% chm & chemvar %in% val | var %in% 'temp' & chemvar %in% tmp) %>%
    filter(station %in% sta) %>% 
    select(-chemvar) %>% 
    spread(var, chemval)

  p <- ggplot(dat, aes_string(x = 'temp', y = chm, size = 'dissval')) +
    geom_point(aes(fill = mo, group = station), pch = 21) +
    stat_smooth(method = 'lm', se = F, colour = 'red') +
    facet_grid(~ cohortyr) +
    ylab(chmlab) +
    xlab(paste(tmp, 'temp')) +
    scale_size(dis, range = c(2, 10)) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      # legend.title = element_blank(), 
      legend.position = 'top'
    ) + 
    guides(
      fill = guide_legend('', override.aes = list(size = 4)),
      size = guide_legend(override.aes = list(linetype = 0))
      )

  return(p)
  
}, height = 315, width = 850)

```

The third set of plots show the selected environmental measure against oxygen with points sized by the pteropod response measure. The x-axis is alway oxygen and can be selected to show minimum, average, and maximum. Plots are facetted by cohort-year with points in each facet colored by month. 
```{r}
column(12, 
  column(4, 
    selectInput('oxy', 'Select x-axis oxygen value:', choices = c('min', 'ave', 'max', 'std', 'rng', 'dlt'))
  )
)

oxydat <- reactive({
  
  # input
  chm <- input$chm
  dis <- input$dis
  val <- input$val
  sta <- input$sta
  oxy <- input$oxy
  chmlab <- chmlab()
  
  dat <- chmdatsum %>% 
    filter(var %in% c(chm, 'oxy')) %>% 
    full_join(biodat, by = c('station', 'date', 'yr', 'cohortyr', 'mo')) %>% 
    select(-lon, -lat, -abu, -avesz, -stdsz) %>% 
    gather('dissvar', 'dissval', typ1:typ3) %>% 
    gather('chemvar', 'chemval', ave:dlt) %>% 
    filter(dissvar %in% dis) %>% 
    filter(station %in% sta) %>% 
    filter(var %in% chm & chemvar %in% val | var %in% 'oxy' & chemvar %in% oxy) %>% 
    select(-chemvar) %>% 
    spread(var, chemval)
  
  return(dat)
  
})

lmmod <- reactive({
  
  # input
  oxydat <- oxydat()
  modtyp <- input$modtyp
  
  datjul <- oxydat %>% 
    filter(mo %in% c('Jul'))
  datapr <- oxydat %>% 
    filter(mo %in% c('Apr'))
  datsep <- oxydat %>% 
    filter(mo %in% c('Sep'))

  frm <- paste('dissval ~', input$chm, '+oxy')
  if(modtyp)
    frm <- gsub('\\+', '*', frm)

  modjul <- paste('lm(', frm, ', data = datjul)')
  modjul <- eval(parse(text = modjul))
  
  modsep <- paste('lm(', frm, ', data = datsep)')
  modsep <- eval(parse(text = modsep))
  
  modapr <- paste('lm(', frm, ', data = datapr)')
  modapr <- eval(parse(text = modapr))
  
  return(list(July = modjul, September = modsep, April = modapr))

})

renderPlot({
  
  validate(
    need(input$chm != 'oxy', 'oxygen on both plot axes not possible')
  )
  
  # inputs
  oxydat <- oxydat()
  chm <- input$chm
  dis <- input$dis
  oxy <- input$oxy
  chmlab <- chmlab()

  p <- ggplot(oxydat, aes_string(x = 'oxy', y = chm, size = 'dissval')) +
    geom_point(aes(fill = mo, group = station), pch = 21) +
    stat_smooth(aes(colour = mo), method = 'lm', se = F, size = 0.7) +
    stat_smooth(method = 'lm', se = F, colour = 'red') +
    facet_grid(~ cohortyr) +
    ylab(chmlab) +
    xlab(paste(oxy, 'oxy')) +
    scale_size(dis, range = c(2, 10)) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      # legend.title = element_blank(), 
      legend.position = 'top'
    ) + 
    guides(
      fill = guide_legend('', override.aes = list(size = 4)),
      size = guide_legend(override.aes = list(linetype = 0))
      )

  return(p)
  
}, height = 315, width = 850)
```

#### {.tabset .tabset-pills}

```{r}
column(12, 
column(width = 4, 
       selectInput('modtyp', "Include interaction:", choices = c(T, F))
       )
)
```


##### July model

```{r}
renderPrint({
  
  validate(
    need(input$chm != 'oxy', 'model not possible with selection')
  )
  
  # input
  lmmod <- lmmod()
  
  summary(lmmod$July)
  
})

# july mod
renderPlot({
  
  req(input$chm != 'oxy')
  
  # input
  lmmod <- lmmod()
  chmlab <- chmlab()
  oxy <- input$oxy
  
  lmmodplo <- lmmod$July
  
  # colors
  cols <- RColorBrewer::brewer.pal(9, 'RdBu')
  
  # get data to plot
  pldat <- get_pldat(lmmodplo, 'oxy')
  pl1 <- pldat[[1]]
  pl2 <- pldat[[2]]
  
  # names for aes_string
  nms <- names(pl1)
  p1 <- ggplot() +
    geom_ribbon(data = pl1, aes_string(x = nms[3], ymin = 'lo', ymax = 'hi', group = nms[4]), alpha = 0.5, fill = 'grey') +
    geom_line(data = pl1, aes_string(x = nms[3], y = nms[1], group = nms[4], colour = nms[4]), size = 1) + 
    geom_text(data = pl2, aes(x= x, y = y, label = lab), hjust = 0) +
    theme_bw() +
    scale_x_continuous(chmlab) +
    scale_colour_gradientn(paste(oxy, 'oxy'), colours = cols) + 
    scale_y_continuous(nms[1]) +
    ggtitle('July model')
  
  return(p1)
  
}, height = 300, width = 600)
```

##### September model

```{r}
renderPrint({
  
  validate(
    need(input$chm != 'oxy', 'model not possible with selection')
  )
  
  # input
  lmmod <- lmmod()
  
  summary(lmmod$September)
  
})

# sept mod
renderPlot({
  
  req(input$chm != 'oxy')
  
  # input
  lmmod <- lmmod()
  chmlab <- chmlab()
  oxy <- input$oxy
  
  lmmodplo <- lmmod$September
  
  # colors
  cols <- RColorBrewer::brewer.pal(9, 'RdBu')
  
  # get data to plot
  pldat <- get_pldat(lmmodplo, 'oxy')
  pl1 <- pldat[[1]]
  pl2 <- pldat[[2]]
  
  # names for aes_string
  nms <- names(pl1)
  p1 <- ggplot() +
    geom_ribbon(data = pl1, aes_string(x = nms[3], ymin = 'lo', ymax = 'hi', group = nms[4]), alpha = 0.5, fill = 'grey') +
    geom_line(data = pl1, aes_string(x = nms[3], y = nms[1], group = nms[4], colour = nms[4]), size = 1) + 
    geom_text(data = pl2, aes(x= x, y = y, label = lab), hjust = 0) +
    theme_bw() +
    scale_x_continuous(chmlab) +
    scale_colour_gradientn(paste(oxy, 'oxy'), colours = cols) + 
    scale_y_continuous(nms[1]) +
    ggtitle('September model')
  
  return(p1)
  
}, height = 300, width = 600)
```

##### April model

```{r}
renderPrint({
  
  validate(
    need(input$chm != 'oxy', 'model not possible with selection')
  )
  
  # input
  lmmod <- lmmod()
  
  summary(lmmod$April)
  
})

# april mod
renderPlot({
  
  req(input$chm != 'oxy')
  
  # input
  lmmod <- lmmod()
  chmlab <- chmlab()
  oxy <- input$oxy

  lmmodplo <- lmmod$April

  # colors
  cols <- RColorBrewer::brewer.pal(9, 'RdBu')
  
  # get data to plot
  pldat <- get_pldat(lmmodplo, 'oxy')
  pl1 <- pldat[[1]]
  pl2 <- pldat[[2]]
  
  # names for aes_string
  nms <- names(pl1)
  p1 <- ggplot() +
    geom_ribbon(data = pl1, aes_string(x = nms[3], ymin = 'lo', ymax = 'hi', group = nms[4]), alpha = 0.5, fill = 'grey') +
    geom_line(data = pl1, aes_string(x = nms[3], y = nms[1], group = nms[4], colour = nms[4]), size = 1) + 
    geom_text(data = pl2, aes(x= x, y = y, label = lab), hjust = 0) +
    theme_bw() +
    scale_x_continuous(chmlab) +
    scale_colour_gradientn(paste(oxy, 'oxy'), colours = cols) + 
    scale_y_continuous(nms[1]) +
    ggtitle('April model')
  
  return(p1)
  
}, height = 300, width = 600)

```

### Grouped by season

The first set of plots show the selected pteropod response measure against the selected environmental measure observed at the time of sampling.  Plots are facetted by cohort-year with points in each facet colored by cohort year. 
```{r}
renderPlot({
  
  # inputs
  disdat <- disdat()
  dis <- input$dis
  chmlab <- chmlab()

  p <- ggplot(disdat, aes(x = chemval, y = dissval)) +
    # geom_line(aes(group = station), colour = 'grey') +
    stat_smooth(method = 'lm', se = F, colour = 'red') +
    geom_label(aes(fill = factor(cohortyr), group = station, label = station), colour = 'grey', size = 3) +
    facet_grid(~ mo) + #, scales = 'free') +
    ylab(dis) +
    xlab(chmlab) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      legend.title = element_blank(), 
      legend.position = 'top'
    )

  return(p)
  
}, height = 300, width = 850)

renderPrint({lapply(dismod()[c('Jul', 'Sep', 'Apr')], summary)})
```

The second set of plots show the selected environmental measure against temperature with points sized by the pteropod response measure. The x-axis is alway temperature and can be selected to show minimum, average, and maximum. Plots are facetted by cohort-year with points in each facet colored by cohort year. 
```{r}
column(12, 
  column(4, 
    selectInput('tmp2', 'Select x-axis temperature value:', choices = c('min', 'ave', 'max', 'std', 'rng', 'dlt'))
  )
)

renderPlot({
  
  validate(
    need(input$chm != 'temp', 'temperature on both plot axes not possible')
  )
  
  # inputs
  chm <- input$chm
  dis <- input$dis
  val <- input$val
  sta <- input$sta
  tmp2 <- input$tmp2
  chmlab <- chmlab()
  
  dat <- chmdatsum %>% 
    filter(var %in% c(chm, 'temp')) %>% 
    full_join(biodat, by = c('station', 'date', 'yr', 'cohortyr', 'mo')) %>% 
    select(-lon, -lat, -abu, -avesz, -stdsz) %>% 
    gather('dissvar', 'dissval', typ1:typ3) %>% 
    gather('chemvar', 'chemval', ave:dlt) %>% 
    filter(dissvar %in% dis) %>% 
    filter(station %in% sta) %>% 
    filter(var %in% chm & chemvar %in% val | var %in% 'temp' & chemvar %in% tmp2) %>% 
    select(-chemvar) %>% 
    spread(var, chemval)

  p <- ggplot(dat, aes_string(x = 'temp', y = chm, size = 'dissval')) +
    geom_point(aes(fill = factor(cohortyr), group = station), pch = 21) +
    stat_smooth(method = 'lm', se = F, colour = 'red') +
    facet_grid(~ mo) +
    ylab(chmlab) +
    xlab(paste(tmp2, 'temp')) +
    scale_size(dis, range = c(2, 10)) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      # legend.title = element_blank(), 
      legend.position = 'top'
    ) + 
    guides(
      fill = guide_legend('', override.aes = list(size = 4)),
      size = guide_legend(override.aes = list(linetype = 0))
      )

  return(p)
  
}, height = 315, width = 850)

```

The third set of plots show the selected environmental measure against oxygen with points sized by the pteropod response measure. The x-axis is alway oxygen and can be selected to show minimum, average, and maximum. Plots are facetted by cohort-year with points in each facet colored by cohort year. 
```{r}
column(12, 
  column(4, 
    selectInput('oxy2', 'Select x-axis temperature value:', choices = c('min', 'ave', 'max', 'std', 'rng', 'dlt'))
  )
)

oxydat2 <- reactive({
  
  # input
  chm <- input$chm
  dis <- input$dis
  val <- input$val
  sta <- input$sta
  oxy2 <- input$oxy2
  chmlab <- chmlab()
  
  dat <- chmdatsum %>% 
    filter(var %in% c(chm, 'oxy')) %>% 
    full_join(biodat, by = c('station', 'date', 'yr', 'cohortyr', 'mo')) %>% 
    select(-lon, -lat, -abu, -avesz, -stdsz) %>% 
    gather('dissvar', 'dissval', typ1:typ3) %>% 
    gather('chemvar', 'chemval', ave:dlt) %>% 
    filter(dissvar %in% dis) %>% 
    filter(station %in% sta) %>% 
    filter(var %in% chm & chemvar %in% val | var %in% 'oxy' & chemvar %in% oxy2) %>% 
    select(-chemvar) %>% 
    spread(var, chemval)
  
  return(dat)
  
})

renderPlot({
  
  validate(
    need(input$chm != 'oxy', 'oxygen on both plot axes not possible')
  )
  
  # inputs
  oxydat2 <- oxydat2()
  chm <- input$chm
  dis <- input$dis
  oxy2 <- input$oxy2
  chmlab <- chmlab()
  
  p <- ggplot(oxydat2, aes_string(x = 'oxy', y = chm, size = 'dissval')) +
    geom_point(aes(fill = factor(cohortyr), group = station), pch = 21) +
    stat_smooth(aes(colour = cohortyr), method = 'lm', se = F, size = 0.7) +
    stat_smooth(method = 'lm', se = F, colour = 'red') +
    facet_grid(~ mo) +
    ylab(chmlab) +
    xlab(paste(oxy2, 'oxy')) +
    scale_size(dis, range = c(2, 10)) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      # legend.title = element_blank(), 
      legend.position = 'top'
    ) + 
    guides(
      fill = guide_legend('', override.aes = list(size = 4)),
      size = guide_legend(override.aes = list(linetype = 0))
      )

  return(p)
  
}, height = 315, width = 850)

```


### By station and cumulative stress {.tabset .tabset-pills}

```{r}
column(12, 
       
       column(4, numericInput('thrsh', 'Under-saturation threshold:', value = 1, min = 0, max = 5, step = 0.05))
       
       )
```

#### By station, profile plots

```{r}
column(12, 
       
       column(4, selectInput('frscl', 'Free y-scale:', choices = c(TRUE, FALSE)))
       
       )
renderPlot({
  
  validate(
    need(input$chm != 'araund', 'profile plots not appropriate for araund')
  )
  
  # input 
  chm <- input$chm
  sta <- input$sta
  frscl <- input$frscl
  
  # date levels, labels
  dtlev <- as.character(sort(unique(chmdatraw$date)))
  dtlab <- as.character(format(unique(chmdatraw$date), '%Y-%m'))

  # chemistry raw data
  toplo <- chmdatraw %>% 
    filter(var %in% chm) %>% 
    filter(station %in% sta) %>% 
    mutate(
      date = factor(date, levels = dtlev, labels = dtlab)
    ) %>% 
    filter(!is.na(val)) %>% 
    group_by(date, mo, station, depth, var) %>% 
    summarise(val = mean(val))

  p1 <- ggplot(toplo, aes(x = val, y = -depth)) +
      geom_path() + 
      geom_point(aes(fill = mo), pch = 21, colour = 'black', size = 4) + 
      ylab('Depth') +
      xlab(chm) + 
      theme_bw(base_family = 'serif', base_size = 14) +
      theme(
        strip.background = element_blank(), 
        legend.title = element_blank(),
        # axis.text.x  = element_text(size = 10, angle = 45, hjust = 1),
        legend.position = 'top'
        ) 
  
  if(frscl)
    p1 <- p1 + 
      facet_grid(station~date, scales = 'free_y')
  else 
    p1 <- p1 + 
      facet_grid(station~date)
  
  return(p1)

}, height = 800, width = 900)
```

#### By station, response and environment

These plots show the selected pteropod response measure and selected environmental measure collected over time with facets for each station.  Points are joined within cohort-years and colored by month. The red line in the right plot shows a saturation state equal to `r renderText({input$thrsh})` when aragonite is selected.

```{r}
renderPlot({
  
  # input 
  dis <- input$dis
  chm <- input$chm
  val <- input$val
  sta <- input$sta
  thrsh <- input$thrsh
  strdat <- strdat()
  chmlab <- chmlab()
  
  # bio
  toplo1 <- biodat %>% 
    select(date, yr, cohortyr, mo, station, typ1, typ2, typ3) %>% 
    gather('diss', 'val', typ1:typ3) %>% 
    filter(diss %in% dis) %>% 
    filter(station %in% sta)
  
  p1 <- ggplot(toplo1, aes(x = factor(date), y = val)) +
    geom_line(aes(group = cohortyr)) +
    geom_point(aes(fill = mo), alpha = 0.7, size = 4, pch = 21) +
    facet_grid(station~.) +
    theme_bw(base_family = 'serif', base_size = 14) +
    theme(
      strip.background = element_blank(), 
      axis.title.x = element_blank(), 
      legend.title = element_blank(),
      axis.text.x  = element_text(size = 10, angle = 45, hjust = 1),
      # panel.grid = element_blank(), 
      legend.position = 'top'
      ) +
    ylab(dis) + 
    scale_x_discrete(labels = format(sort(unique(toplo1$date)), '%Y-%m'))
  
  # chem
  toplo2 <- chmdatsum %>% 
    filter(var %in% chm) %>% 
    gather('estv', 'valv', ave:dlt) %>% 
    filter(estv %in% val) %>% 
    filter(station %in% sta)
  
  p2 <- ggplot(toplo2, aes(x = factor(date), y = valv)) +
    geom_line(aes(group = cohortyr)) +
    geom_point(aes(fill = mo), alpha = 0.7, size  = 4, pch = 21) +
    facet_grid(station~.) +
    theme_bw(base_family = 'serif', base_size = 14) + 
    theme(
      strip.background = element_blank(), 
      axis.title.x = element_blank(), 
      legend.title = element_blank(),
      axis.text.x  = element_text(size = 10, angle = 45, hjust = 1),
      # panel.grid = element_blank(), 
      legend.position = 'top'
      ) +
    ylab(chmlab) + 
    scale_x_discrete(labels = format(sort(unique(toplo2$date)), '%Y-%m'))

  # add 1 horizontal or aragonite
  if(chm == 'ara')
    p2 <- p2 + 
      geom_hline(yintercept = thrsh, colour = 'tomato1', size = 1, alpha = 0.7) +
      geom_text(data = strdat, aes(y = thrsh, label = paste(strsdis, round(strscnt, 1), sep = ', ')), colour = 'tomato1', vjust = 1, hjust = 0)
  
  grid.arrange(p1, p2, ncol = 2)
  
}, height = 900, width = 900)
```

#### Evaluating cumulative stress, discrete

Within each cohort year, the "cumulative stress" was estimated as the duration of time within a cohort year when under-saturation of aragonite occcurred.  For "discrete stress", the number of times a sampling period was under-saturated was counted for all of the sampling months in a cohort-year.  This produced an estimate of total stress within a cohort-year as the cumulative measure from zero to three, where zero meant that none of the sampling periods were under-saturated and three meant all were under-saturated.  The cumulative stress counts assigned to each station and cohort year can be viewed in the first tab.

```{r}

renderPlot({
  
  validate(
    need(input$chm == 'ara', 'plots only valid for aragonite')
  )
  
  # input
  strdat <- strdat()
  dis <- input$dis

  # rename pteropod response measure to generic
  toplo <- strdat %>% 
    rename(
      rsp = !!dis
    )
  
  p1 <- ggplot(toplo, aes(x = strsdis, y = rsp)) +
    geom_line(aes(group = costa)) +
    # geom_point(aes(fill = mo, group = costa), pch = 21, size = 4) +
    geom_label(aes(fill = mo, group = costa, label = station), size = 3, colour = 'grey') +
    geom_smooth(method = 'lm', colour = 'red') +
    xlab('Cumulative stress magnitude') +
    ylab(dis) +
    facet_wrap(~cohortyr) +
    theme_bw(base_family = 'serif', base_size = 16) + 
    theme(
      legend.position = 'top', 
      legend.title = element_blank(), 
      strip.background = element_blank()
    )
  
  p2 <- ggplot(toplo, aes(x = mo, y = strsdis, group = costa)) +
    geom_line() +
    # geom_jitter(aes(fill = mo, size = rsp), pch = 21, width = 0.1, height = 0.1) +
    geom_label(aes(fill = mo, group = costa, label = station, size = rsp), colour = 'grey', position = position_jitter(width = 0.1, height = 0.1)) +
    ylab('Cumulative stress magnitude') +
    facet_wrap(~cohortyr) +
    theme_bw(base_family = 'serif', base_size = 16) + 
    theme(
      axis.title.x = element_blank(), 
      legend.position = 'top', 
      strip.background = element_blank()
    ) + 
    scale_size(dis) +
    guides(fill = guide_legend(title = element_blank(), override.aes = list(size = 4)))
  
  p1 + p2 + plot_layout(ncol = 1, heights = c(1, 1))

}, height = 600, width = 850)
  
```

#### Evaluating cumulative stress, continuous

Within each cohort year, the "cumulative stress" was estimated as the duration of time within a cohort year when under-saturation of aragonite occcurred.  For "continuous stress", the difference between measured aragonite saturation and the selected under-saturation was used as measure of cumulative stress in a cohort-year.  This produced an estimate of total stress within a cohort-year as the total sum of the difference.  The cumulative stress counts assigned to each station and cohort year can be viewed in the first tab.

```{r}
renderPlot({
  
  validate(
    need(input$chm == 'ara', 'plots only valid for aragonite')
  )
  
  # input
  strdat <- strdat()
  dis <- input$dis

  # rename pteropod response measure to generic
  toplo <- strdat %>% 
    rename(
      rsp = !!dis
    )
  
  p1 <- ggplot(toplo, aes(x = strscnt, y = rsp)) +
    geom_line(aes(group = costa)) +
    # geom_point(aes(fill = mo, group = costa), pch = 21, size = 4) +
    geom_label(aes(fill = mo, group = costa, label = station), size = 3, colour = 'grey') +
    geom_smooth(method = 'lm', colour = 'red') +
    xlab('Cumulative stress magnitude') +
    ylab(dis) +
    facet_wrap(~cohortyr) +
    theme_bw(base_family = 'serif', base_size = 16) + 
    theme(
      legend.position = 'top', 
      legend.title = element_blank(), 
      strip.background = element_blank()
    )
  
  p2 <- ggplot(toplo, aes(x = mo, y = strscnt, group = costa)) +
    geom_line() +
    # geom_jitter(aes(fill = mo, size = rsp), pch = 21, width = 0.1, height = 0.1) +
    geom_label(aes(fill = mo, group = costa, label = station, size = rsp), colour = 'grey', position = position_jitter(width = 0.1, height = 0.1)) +
    ylab('Cumulative stress magnitude') +
    facet_wrap(~cohortyr) +
    theme_bw(base_family = 'serif', base_size = 16) + 
    theme(
      axis.title.x = element_blank(), 
      legend.position = 'top', 
      strip.background = element_blank()
    ) + 
    scale_size(dis) +
    guides(fill = guide_legend(title = element_blank(), override.aes = list(size = 4)))
  
  p1 + p2 + plot_layout(ncol = 1, heights = c(1, 1))

}, height = 600, width = 850)
  
```

### Growth {.tabset .tabset-pills}

```{r}
grodat <- reactive({

  # inputs
  chm <- input$chm
  val <- input$val
  sta <- input$sta

  dat <- chmdatsum %>% 
    filter(var %in% chm) %>% 
    full_join(biodat, by = c('station', 'date', 'yr', 'cohortyr', 'mo')) %>% 
    select(-lon, -lat, -var, -abu, -typ1, -typ2, -typ3, -stdsz, -`% type II&III`) %>%
    gather('chemvar', 'chemval', ave:dlt) %>% 
    filter(chemvar %in% val) %>% 
    filter(station %in% sta)
  
  return(dat)
  
})

gromod <- reactive({
  
  # input
  grodat <- grodat()
  
  # complete model
  all <- lm(avesz ~ chemval, data = grodat)
  
  # by cohort years
  bycohortyr <- grodat %>% 
    group_by(cohortyr) %>% 
    nest %>% 
    mutate(
      mod = purrr::map(data, function(x){ try({lm(avesz ~ chemval, data = x)})})
    )
  bycohortyrout <- pull(bycohortyr, mod)
  names(bycohortyrout) <- pull(bycohortyr, cohortyr)
  
  # by seasons
  byseas <- grodat %>% 
    group_by(mo) %>% 
    nest %>% 
    mutate(
      mod = purrr::map(data, function(x){ try({lm(avesz ~ chemval, data = x)})})
    )
  byseasout <- pull(byseas, mod)
  names(byseasout) <- pull(byseas, mo)

  # combine all    
  out <- c(
    all = list(all), 
    bycohortyrout,
    byseasout
  )
  
  return(out)
  
})
```

```{r}
renderPlot({
  
  # inputs
  grodat <- grodat()
  chmlab <- chmlab()

  p <- ggplot(grodat, aes(x = chemval, y = avesz)) +
    # geom_line(aes(group = station), colour = 'grey') +
    stat_smooth(method = 'lm', se =T, colour = 'red') +
    geom_label(aes(fill = mo, group = station, label = station), size = 3, colour = 'grey') +
    ylab("Average size") +
    xlab(chmlab) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      legend.title = element_blank(), 
      legend.position = 'top'
    )

  return(p)
  
}, height = 400, width = 400)

renderPrint({lapply(gromod()['all'], summary)})

renderPlot({
  
  # inputs
  grodat <- grodat()
  chmlab <- chmlab()

  p <- ggplot(grodat, aes(x = chemval, y = avesz)) +
    # geom_line(aes(group = station), colour = 'grey') +
    stat_smooth(method = 'lm', se = F, colour = 'red') +
    geom_label(aes(fill = mo, group = station, label = station), size = 3, colour = 'grey') +
    facet_grid(~ cohortyr) + #, scales = 'free') +
    ylab("Average size") +
    xlab(chmlab) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      legend.title = element_blank(), 
      legend.position = 'top'
    )

  return(p)
  
}, height = 300, width = 850)

renderPrint({lapply(gromod()[c('2014', '2015', '2016')], summary)})

renderPlot({
  
  # inputs
  grodat <- grodat()
  chmlab <- chmlab()

  p <- ggplot(grodat, aes(x = chemval, y = avesz)) +
    # geom_line(aes(group = station), colour = 'grey') +
    stat_smooth(method = 'lm', se = F, colour = 'red') +
    geom_label(aes(fill = cohortyr, group = station, label = station), size = 3, colour = 'grey') +
    facet_grid(~ mo) + #, scales = 'free') +
    ylab("Average size") +
    xlab(chmlab) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      legend.title = element_blank(), 
      legend.position = 'top'
    )

  return(p)
  
}, height = 300, width = 850)

renderPrint({lapply(gromod()[c('Jul', 'Sep', 'Apr')], summary)})
```

### Abundance {.tabset .tabset-pills}

```{r}
abudat <- reactive({

  # inputs
  chm <- input$chm
  val <- input$val
  sta <- input$sta

  dat <- chmdatsum %>% 
    filter(var %in% chm) %>% 
    full_join(biodat, by = c('station', 'date', 'yr', 'cohortyr', 'mo')) %>% 
    select(-lon, -lat, -var, -typ1, -typ2, -typ3, -avesz, -stdsz, -`% type II&III`) %>%
    gather('chemvar', 'chemval', ave:dlt) %>% 
    filter(chemvar %in% val) %>% 
    filter(station %in% sta)
  
  return(dat)
  
})

abumod <- reactive({
  
  # input
  abudat <- abudat()
  
  # complete model
  all <- lm(log10(abu) ~ chemval, data = abudat)
  
  # by cohort years
  bycohortyr <- abudat %>% 
    group_by(cohortyr) %>% 
    nest %>% 
    mutate(
      mod = purrr::map(data, function(x){ try({lm(log10(abu) ~ chemval, data = x)})})
    )
  bycohortyrout <- pull(bycohortyr, mod)
  names(bycohortyrout) <- pull(bycohortyr, cohortyr)
  
  # by seasons
  byseas <- abudat %>% 
    group_by(mo) %>% 
    nest %>% 
    mutate(
      mod = purrr::map(data, function(x){ try({lm(log10(abu) ~ chemval, data = x)})})
    )
  byseasout <- pull(byseas, mod)
  names(byseasout) <- pull(byseas, mo)

  # combine all    
  out <- c(
    all = list(all), 
    bycohortyrout,
    byseasout
  )
  
  return(out)
  
})
```

```{r}
renderPlot({
  
  # inputs
  abudat <- abudat()
  chmlab <- chmlab()

  p <- ggplot(abudat, aes(x = chemval, y = abu)) +
    # geom_line(aes(group = station), colour = 'grey') +
    stat_smooth(method = 'lm', se =T, colour = 'red') +
    geom_label(aes(fill = mo, group = station, label = station), size = 3, colour = 'grey') +
    ylab("Average size") +
    xlab(chmlab) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      legend.title = element_blank(), 
      legend.position = 'top'
    ) + 
    scale_y_log10('log10(abu)')

  return(p)
  
}, height = 400, width = 400)

renderPrint({lapply(abumod()['all'], summary)})

renderPlot({
  
  # inputs
  abudat <- abudat()
  chmlab <- chmlab()

  p <- ggplot(abudat, aes(x = chemval, y = abu)) +
    # geom_line(aes(group = station), colour = 'grey') +
    stat_smooth(method = 'lm', se = F, colour = 'red') +
    geom_label(aes(fill = mo, group = station, label = station), size = 3, colour = 'grey') +
    facet_grid(~ cohortyr) + #, scales = 'free') +
    ylab("Average size") +
    xlab(chmlab) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      legend.title = element_blank(), 
      legend.position = 'top'
    ) + 
    scale_y_log10('log10(abu)')

  return(p)
  
}, height = 300, width = 850)

renderPrint({lapply(abumod()[c('2014', '2015', '2016')], summary)})

renderPlot({
  
  # inputs
  abudat <- abudat()
  chmlab <- chmlab()

  p <- ggplot(abudat, aes(x = chemval, y = abu)) +
    # geom_line(aes(group = station), colour = 'grey') +
    stat_smooth(method = 'lm', se = F, colour = 'red') +
    geom_label(aes(fill = cohortyr, group = station, label = station), size = 3, colour = 'grey') +
    facet_grid(~ mo) + #, scales = 'free') +
    ylab("Average size") +
    xlab(chmlab) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      legend.title = element_blank(), 
      legend.position = 'top'
    ) + 
    scale_y_log10('log10(abu)')

  return(p)
  
}, height = 300, width = 850)

renderPrint({lapply(abumod()[c('Jul', 'Sep', 'Apr')], summary)})
```
