---
title: "Pteropod Puget Sound time series"
author: ""
output: 
  html_document
self_contained: yes
runtime: shiny
---

```{r, message = F, warning = F, echo = F}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, dev.args = list(bg = 'transparent'), eval = T, echo = F)

library(tidyverse)
library(shiny)
library(gridExtra)
library(glue)
library(lubridate)
library(patchwork)
library(shinyWidgets)

source('R/funcs.R')

data(chmdatsum)
data(chmdatraw)
data(biodat)

# pteropod birthday
strt <- '2008-06-01' %>% 
  as.Date

# make cohortyr an ordered factor
biodat <- biodat %>% 
  mutate(cohortyr = factor(cohortyr, ordered = T))
chmdatsum <- chmdatsum %>% 
  mutate(cohortyr = factor(cohortyr, ordered = T))
```

```{r reactives}
# cumulative stress data
strdat <- reactive({

  # inputs
  val <- input$val
  sta <- input$sta
  thrsh <- input$thrsh

  out <- chmdatsum %>% 
    filter(var %in% 'ara') %>% 
    full_join(biodat, by = c('station', 'date', 'yr', 'cohortyr', 'mo')) %>% 
    select(-lon, -lat, -var, -abu, -len, -yr) %>% 
    gather('ara', 'chmval', ave:dlt) %>% 
    filter(ara %in% val) %>% 
    filter(station %in% sta) %>% 
    mutate(
      aracat = ifelse(chmval < thrsh, 1, 0),
      aradff = thrsh - chmval
      ) %>% 
    unite('costa', cohortyr, station, remove = F) %>% 
    group_by(costa) %>% 
    mutate(
      strsdis = cumsum(aracat),
      strscnt = cumsum(aradff)
    ) 

  return(out)
  
})

# wq chem labels
chmlab <- reactive({
  
  # input
  chm <- input$chm
  val <- input$val
  
  if(chm == 'araund')
    return(chm)
  
  out <- paste(val, chm)
  
  return(out)
  
})
```

```{r}
column(12, 
  column(4, 
    selectInput('chm', 'Select chemistry:', choices = c('ara', 'araund', 'co3', 'ph', 'pco2', 'temp', 'oxy', 'chla'))
  ),
  
  column(4, 
    selectInput('val', 'Select chemistry value:', choices = c('min', 'ave', 'max', 'std', 'rng', 'dlt'))
  ),
  
  column(4, 
    selectInput('dis', 'Select dissolution:', choices = c('typ1', 'typ2', 'typ3'), selected = 'typ3')
  )
)
column(12, 
  column(4, 
  
    pickerInput(inputId = "sta", label = 'Select stations:', choices = unique(chmdatsum$station),
      options = list(`actions-box` = TRUE, size = 20), selected = unique(chmdatsum$station), multiple = TRUE)    
              
  )       
)
```

Year values shown correspond to "cohort-years" that begin in June with spring spawning, i.e., the April "cohort-year" is one year prior to the calendar year.  

```{r}
# data for first set of plots in first three tabs
disdat <- reactive({
  
  # inputs
  chm <- input$chm
  dis <- input$dis
  val <- input$val
  sta <- input$sta
  
  dat <- chmdatsum %>% 
    filter(var %in% chm) %>% 
    full_join(biodat, by = c('station', 'date', 'yr', 'cohortyr', 'mo')) %>% 
    select(-lon, -lat, -var, -abu, -len) %>% 
    gather('dissvar', 'dissval', typ1:typ3) %>% 
    gather('chemvar', 'chemval', ave:dlt) %>% 
    filter(dissvar %in% dis) %>% 
    filter(chemvar %in% val) %>% 
    filter(station %in% sta)  
  
  return(dat)
  
})

# regression mods for first set of plots in first three tabs
dismod <- reactive({
  
  # input
  disdat <- disdat()
  
  # complete model
  all <- lm(dissval ~ chemval, data = disdat)
  
  # by cohort years
  bycohortyr <- disdat %>% 
    group_by(cohortyr) %>% 
    nest %>% 
    mutate(
      mod = purrr::map(data, function(x){ try({lm(dissval ~ chemval, data = x)})})
    )
  bycohortyrout <- pull(bycohortyr, mod)
  names(bycohortyrout) <- pull(bycohortyr, cohortyr)

  # by seasons
  byseas <- disdat %>% 
    group_by(mo) %>% 
    nest %>% 
    mutate(
      mod = purrr::map(data, function(x){ try({lm(dissval ~ chemval, data = x)})})
    )
  byseasout <- pull(byseas, mod)
  names(byseasout) <- pull(byseas, mo)

  # combine all    
  out <- c(
    all = list(all), 
    bycohortyrout,
    byseasout
    )
  
  return(out)
  
})
```

## {.tabset}

### All data {.tabset .tabset-pills}

#### Response vs one variable

This plot shows the selected pteropod response measure against the selected environmental measure observed at the time of sampling.

```{r}
renderPlot({
  
  # inputs
  disdat <- disdat()
  dis <- input$dis
  chmlab <- chmlab()

  p <- ggplot(disdat, aes(x = chemval, y = dissval)) +
    # geom_line(aes(group = station), colour = 'grey') +
    stat_smooth(method = 'lm', se = T, colour = 'red') +
    geom_label(aes(fill = mo, group = station, label = station), colour = 'grey', size = 3) +
    ylab(dis) +
    xlab(chmlab) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      legend.title = element_blank(), 
      legend.position = 'top'
    )

  return(p)
  
}, height = 400, width = 400)

renderPrint({lapply(dismod()['all'], summary)})
```

#### Response vs two variables

```{r}
column(12, 
       column(4, 
              selectInput('chm2', 'Select second chemistry variable:', choices = c('ara', 'araund', 'co3', 'ph', 'pco2', 'temp', 'oxy', 'chla'), selected = 'temp')
       ),
       
       column(4, 
              selectInput('val2', 'Select second chemistry value:', choices = c('min', 'ave', 'max', 'std', 'rng', 'dlt'))
       ),
       column(width = 4, 
              selectInput('modtyptmp', "Include interaction:", choices = c(T, F))
       )
)
column(12, 
       column(4, 
              selectInput('flt', 'Filter data by:', choices = c('nothing', 'Jul', 'Apr', 'Sep', '2014', '2015', '2016'))
              )
       )
```

```{r}
mod2dat <- reactive({  
  
  # inputs
  chm <- input$chm
  dis <- input$dis
  val <- input$val
  chm2 <- input$chm2
  val2 <- input$val2
  sta <- input$sta
  flt <- input$flt
  
  validate(
    need(chm != chm2, 'Variable selections must be different')
  )
  
  dat <- chmdatsum %>% 
    filter(var %in% c(chm, chm2)) %>% 
    full_join(biodat, by = c('station', 'date', 'yr', 'cohortyr', 'mo')) %>% 
    select(-lon, -lat, -abu, -len) %>% 
    gather('dissvar', 'dissval', typ1:typ3) %>% 
    gather('chemvar', 'chemval', ave:dlt) %>% 
    filter(dissvar %in% dis) %>% 
    filter(var %in% chm & chemvar %in% val | var %in% chm2 & chemvar %in% val2) %>%
    filter(station %in% sta) %>% 
    select(-chemvar) %>% 
    spread(var, chemval)

  if(flt %in% c('Jul', 'Sep', 'Apr'))
    dat <- dat %>% filter(mo %in% flt)

  if(flt %in% c('2014', '2015', '2016'))
    dat <- dat %>% filter(cohortyr %in% as.numeric(flt))

  return(dat)
  
})

lmmod2 <- reactive({
  
  # input
  mod2dat <- mod2dat()
  modtyptmp <- input$modtyptmp
  chm <- input$chm
  chm2 <- input$chm2
  
  frm <- paste('dissval ~', chm, '+', chm2)
  if(modtyptmp)
    frm <- gsub('\\+', '*', frm)
  
  mod <- paste('lm(', frm, ', data = mod2dat)')
  mod <- eval(parse(text = mod))
  
  return(mod)
  
})

vifval <- reactive({
  
  # input
  lmmod2<- lmmod2()
  chm2 <- input$chm2
  
  datin <- lmmod2$model[,-1] 
  vifval <- paste0('lm(', chm2, ' ~ ., data= datin)')
  vifval <- eval(parse(text = vifval)) %>% 
    summary %>% 
    .$r.squared
  vifval <- 1 / (1 - vifval)
  vifval <- round(vifval, 2) %>% 
    paste('VIF:', .)
  
  return(vifval)
  
})
 
# interaction plot
output$int2plt <- renderPlot({
  
  # input
  lmmod2 <- lmmod2()
  chmlab <- chmlab()
  chm2 <- input$chm2
  val2 <- input$val2
  dis <- input$dis
  
  # colors
  cols <- RColorBrewer::brewer.pal(9, 'RdBu')
  
  # get data to plot
  pldat <- get_pldat(lmmod2, chm2)
  pl1 <- pldat[[1]]
  pl2 <- pldat[[2]]
  
  # names for aes_string
  nms <- names(pl1)
  p1 <- ggplot() +
    geom_ribbon(data = pl1, aes_string(x = nms[3], ymin = 'lo', ymax = 'hi', group = nms[4]), alpha = 0.5, fill = 'grey') +
    geom_line(data = pl1, aes_string(x = nms[3], y = nms[1], group = nms[4], colour = nms[4]), size = 1) + 
    # geom_text(data = pl2, aes(x= x, y = y, label = lab), hjust = 0) +
    theme_bw() +
    scale_x_continuous(chmlab) +
    scale_colour_gradientn(paste(val2, chm2), colours = cols) + 
    scale_y_continuous(dis)
  
  return(p1)
  
}, height = 300, width = 600) 

```

```{r}
plotOutput('int2plt')
```
<br><br><br>
```{r}
renderText({vifval()})
renderPrint({summary(lmmod2())})
```

### Grouped by year

These plots show the selected pteropod response measure against the selected environmental measure observed at the time of sampling.  Plots are facetted by cohort-year with points in each facet colored by month. 
```{r}
renderPlot({
  
  # inputs
  disdat <- disdat()
  dis <- input$dis
  chmlab <- chmlab()

  p <- ggplot(disdat, aes(x = chemval, y = dissval)) +
    # geom_line(aes(group = station), colour = 'grey') +
    stat_smooth(method = 'lm', se = F, colour = 'red') +
    geom_label(aes(fill = mo, group = station, label = station), colour = 'grey', size = 3) +
    facet_grid(~ cohortyr) + #, scales = 'free') +
    ylab(dis) +
    xlab(chmlab) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      legend.title = element_blank(), 
      legend.position = 'top'
    )

  return(p)
  
}, height = 300, width = 850)

renderPrint({lapply(dismod()[c('2014', '2015', '2016')], summary)})
```

### Grouped by season

These plots show the selected pteropod response measure against the selected environmental measure observed at the time of sampling.  Plots are facetted by cohort-year with points in each facet colored by cohort year. 
```{r}
renderPlot({
  
  # inputs
  disdat <- disdat()
  dis <- input$dis
  chmlab <- chmlab()

  p <- ggplot(disdat, aes(x = chemval, y = dissval)) +
    # geom_line(aes(group = station), colour = 'grey') +
    stat_smooth(method = 'lm', se = F, colour = 'red') +
    geom_label(aes(fill = factor(cohortyr), group = station, label = station), colour = 'grey', size = 3) +
    facet_grid(~ mo) + #, scales = 'free') +
    ylab(dis) +
    xlab(chmlab) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      legend.title = element_blank(), 
      legend.position = 'top'
    )

  return(p)
  
}, height = 300, width = 850)

renderPrint({lapply(dismod()[c('Jul', 'Sep', 'Apr')], summary)})
```

### By station and cumulative stress {.tabset .tabset-pills}

```{r}
column(12, 
       
       column(4, numericInput('thrsh', 'Under-saturation threshold:', value = 1, min = 0, max = 5, step = 0.05))
       
       )
```

#### By station, profile plots

```{r}
column(12, 
       
       column(4, selectInput('frscl', 'Free y-scale:', choices = c(TRUE, FALSE)))
       
       )
renderPlot({
  
  validate(
    need(input$chm != 'araund', 'profile plots not appropriate for araund')
  )
  
  # input 
  chm <- input$chm
  sta <- input$sta
  frscl <- input$frscl
  
  # date levels, labels
  dtlev <- as.character(sort(unique(chmdatraw$date)))
  dtlab <- as.character(format(unique(chmdatraw$date), '%Y-%m'))

  # chemistry raw data
  toplo <- chmdatraw %>% 
    filter(var %in% chm) %>% 
    filter(station %in% sta) %>% 
    mutate(
      date = factor(date, levels = dtlev, labels = dtlab)
    ) %>% 
    filter(!is.na(val)) %>% 
    group_by(date, mo, station, depth, var) %>% 
    summarise(val = mean(val))

  p1 <- ggplot(toplo, aes(x = val, y = -depth)) +
      geom_path() + 
      geom_point(aes(fill = mo), pch = 21, colour = 'black', size = 4) + 
      ylab('Depth') +
      xlab(chm) + 
      theme_bw(base_family = 'serif', base_size = 14) +
      theme(
        strip.background = element_blank(), 
        legend.title = element_blank(),
        # axis.text.x  = element_text(size = 10, angle = 45, hjust = 1),
        legend.position = 'top'
        ) 
  
  if(frscl)
    p1 <- p1 + 
      facet_grid(station~date, scales = 'free_y')
  else 
    p1 <- p1 + 
      facet_grid(station~date)
  
  return(p1)

}, height = 800, width = 900)
```

#### By station, response and environment

These plots show the selected pteropod response measure and selected environmental measure collected over time with facets for each station.  Points are joined within cohort-years and colored by month. The red line in the right plot shows a saturation state equal to `r renderText({input$thrsh})` when aragonite is selected.

```{r}
renderPlot({
  
  # input 
  dis <- input$dis
  chm <- input$chm
  val <- input$val
  sta <- input$sta
  thrsh <- input$thrsh
  strdat <- strdat()
  chmlab <- chmlab()
  
  # bio
  toplo1 <- biodat %>% 
    select(date, yr, cohortyr, mo, station, typ1, typ2, typ3) %>% 
    gather('diss', 'val', typ1:typ3) %>% 
    filter(diss %in% dis) %>% 
    filter(station %in% sta)
  
  p1 <- ggplot(toplo1, aes(x = factor(date), y = val)) +
    geom_line(aes(group = cohortyr)) +
    geom_point(aes(fill = mo), alpha = 0.7, size = 4, pch = 21) +
    facet_grid(station~.) +
    theme_bw(base_family = 'serif', base_size = 14) +
    theme(
      strip.background = element_blank(), 
      axis.title.x = element_blank(), 
      legend.title = element_blank(),
      axis.text.x  = element_text(size = 10, angle = 45, hjust = 1),
      # panel.grid = element_blank(), 
      legend.position = 'top'
      ) +
    ylab(dis) + 
    scale_x_discrete(labels = format(sort(unique(toplo1$date)), '%Y-%m'))
  
  # chem
  toplo2 <- chmdatsum %>% 
    filter(var %in% chm) %>% 
    gather('estv', 'valv', ave:dlt) %>% 
    filter(estv %in% val) %>% 
    filter(station %in% sta)
  
  p2 <- ggplot(toplo2, aes(x = factor(date), y = valv)) +
    geom_line(aes(group = cohortyr)) +
    geom_point(aes(fill = mo), alpha = 0.7, size  = 4, pch = 21) +
    facet_grid(station~.) +
    theme_bw(base_family = 'serif', base_size = 14) + 
    theme(
      strip.background = element_blank(), 
      axis.title.x = element_blank(), 
      legend.title = element_blank(),
      axis.text.x  = element_text(size = 10, angle = 45, hjust = 1),
      # panel.grid = element_blank(), 
      legend.position = 'top'
      ) +
    ylab(chmlab) + 
    scale_x_discrete(labels = format(sort(unique(toplo2$date)), '%Y-%m'))

  # add 1 horizontal or aragonite
  if(chm == 'ara')
    p2 <- p2 + 
      geom_hline(yintercept = thrsh, colour = 'tomato1', size = 1, alpha = 0.7) +
      geom_text(data = strdat, aes(y = thrsh, label = paste(strsdis, round(strscnt, 1), sep = ', ')), colour = 'tomato1', vjust = 1, hjust = 0)
  
  grid.arrange(p1, p2, ncol = 2)
  
}, height = 900, width = 900)
```

#### Evaluating cumulative stress, discrete

Within each cohort year, the "cumulative stress" was estimated as the duration of time within a cohort year when under-saturation of aragonite occcurred.  For "discrete stress", the number of times a sampling period was under-saturated was counted for all of the sampling months in a cohort-year.  This produced an estimate of total stress within a cohort-year as the cumulative measure from zero to three, where zero meant that none of the sampling periods were under-saturated and three meant all were under-saturated.  The cumulative stress counts assigned to each station and cohort year can be viewed in the first tab.

```{r}

renderPlot({
  
  validate(
    need(input$chm == 'ara', 'plots only valid for aragonite')
  )
  
  # input
  strdat <- strdat()
  dis <- input$dis

  # rename pteropod response measure to generic
  toplo <- strdat %>% 
    rename(
      rsp = !!dis
    )
  
  p1 <- ggplot(toplo, aes(x = strsdis, y = rsp)) +
    geom_line(aes(group = costa)) +
    # geom_point(aes(fill = mo, group = costa), pch = 21, size = 4) +
    geom_label(aes(fill = mo, group = costa, label = station), size = 3, colour = 'grey') +
    geom_smooth(method = 'lm', colour = 'red') +
    xlab('Cumulative stress magnitude') +
    ylab(dis) +
    facet_wrap(~cohortyr) +
    theme_bw(base_family = 'serif', base_size = 16) + 
    theme(
      legend.position = 'top', 
      legend.title = element_blank(), 
      strip.background = element_blank()
    )
  
  p2 <- ggplot(toplo, aes(x = mo, y = strsdis, group = costa)) +
    geom_line() +
    # geom_jitter(aes(fill = mo, size = rsp), pch = 21, width = 0.1, height = 0.1) +
    geom_label(aes(fill = mo, group = costa, label = station, size = rsp), colour = 'grey', position = position_jitter(width = 0.1, height = 0.1)) +
    ylab('Cumulative stress magnitude') +
    facet_wrap(~cohortyr) +
    theme_bw(base_family = 'serif', base_size = 16) + 
    theme(
      axis.title.x = element_blank(), 
      legend.position = 'top', 
      strip.background = element_blank()
    ) + 
    scale_size(dis) +
    guides(fill = guide_legend(title = element_blank(), override.aes = list(size = 4)))
  
  p1 + p2 + plot_layout(ncol = 1, heights = c(1, 1))

}, height = 600, width = 850)
  
```

#### Evaluating cumulative stress, continuous

Within each cohort year, the "cumulative stress" was estimated as the duration of time within a cohort year when under-saturation of aragonite occcurred.  For "continuous stress", the difference between measured aragonite saturation and the selected under-saturation was used as measure of cumulative stress in a cohort-year.  This produced an estimate of total stress within a cohort-year as the total sum of the difference.  The cumulative stress counts assigned to each station and cohort year can be viewed in the first tab.

```{r}
renderPlot({
  
  validate(
    need(input$chm == 'ara', 'plots only valid for aragonite')
  )
  
  # input
  strdat <- strdat()
  dis <- input$dis

  # rename pteropod response measure to generic
  toplo <- strdat %>% 
    rename(
      rsp = !!dis
    )
  
  p1 <- ggplot(toplo, aes(x = strscnt, y = rsp)) +
    geom_line(aes(group = costa)) +
    # geom_point(aes(fill = mo, group = costa), pch = 21, size = 4) +
    geom_label(aes(fill = mo, group = costa, label = station), size = 3, colour = 'grey') +
    geom_smooth(method = 'lm', colour = 'red') +
    xlab('Cumulative stress magnitude') +
    ylab(dis) +
    facet_wrap(~cohortyr) +
    theme_bw(base_family = 'serif', base_size = 16) + 
    theme(
      legend.position = 'top', 
      legend.title = element_blank(), 
      strip.background = element_blank()
    )
  
  p2 <- ggplot(toplo, aes(x = mo, y = strscnt, group = costa)) +
    geom_line() +
    # geom_jitter(aes(fill = mo, size = rsp), pch = 21, width = 0.1, height = 0.1) +
    geom_label(aes(fill = mo, group = costa, label = station, size = rsp), colour = 'grey', position = position_jitter(width = 0.1, height = 0.1)) +
    ylab('Cumulative stress magnitude') +
    facet_wrap(~cohortyr) +
    theme_bw(base_family = 'serif', base_size = 16) + 
    theme(
      axis.title.x = element_blank(), 
      legend.position = 'top', 
      strip.background = element_blank()
    ) + 
    scale_size(dis) +
    guides(fill = guide_legend(title = element_blank(), override.aes = list(size = 4)))
  
  p1 + p2 + plot_layout(ncol = 1, heights = c(1, 1))

}, height = 600, width = 850)
  
```

### Growth by environment {.tabset .tabset-pills}

```{r}
grodat_winter <- reactive({

  # inputs
  chm <- input$chm
  val <- input$val
  sta <- input$sta
  dis <- input$dis

  dat <- chmdatsum %>% 
    filter(var %in% chm) %>% 
    full_join(biodat, by = c('station', 'date', 'yr', 'cohortyr', 'mo')) %>% 
    select(date, yr, cohortyr, mo, station, len, !!dis, ave:dlt, cohortseas) %>% 
    filter(cohortseas %in% 'winter') %>% 
    gather('chemvar', 'chemval', ave:dlt) %>% 
    filter(chemvar %in% val) %>% 
    filter(station %in% sta) %>% 
    rename(disval = !!dis)
  
  return(dat)
  
})

gromod_winter <- reactive({
  
  # input
  grodat_winter <- grodat_winter()

  # complete model
  allchm <- lm(len ~ chemval, data = grodat_winter)
  alldis <- lm(len ~ disval, data = grodat_winter)
  

  # combine all    
  out <- c(
    allchm = list(allchm), 
    alldis = list(alldis)
  )
  
  return(out)
  
})

grodat_summer <- reactive({

  # inputs
  chm <- input$chm
  val <- input$val
  sta <- input$sta
  dis <- input$dis

  dat <- chmdatsum %>% 
    filter(var %in% chm) %>% 
    full_join(biodat, by = c('station', 'date', 'yr', 'cohortyr', 'mo')) %>% 
    select(date, yr, cohortyr, mo, station, len, !!dis, ave:dlt, cohortseas) %>% 
    filter(cohortseas %in% 'summer') %>% 
    gather('chemvar', 'chemval', ave:dlt) %>% 
    filter(chemvar %in% val) %>% 
    filter(station %in% sta) %>% 
    rename(disval = !!dis)
  
  return(dat)
  
})

gromod_summer <- reactive({
  
  # input
  grodat_summer <- grodat_summer()

  # complete model
  allchm <- lm(len ~ chemval, data = grodat_summer)
  alldis <- lm(len ~ disval, data = grodat_summer)
  

  # combine all    
  out <- c(
    allchm = list(allchm), 
    alldis = list(alldis)
  )
  
  return(out)
  
})
```

#### Winter cohort

```{r}
renderPlot({
  
  # inputs
  grodat_winter <- grodat_winter()
  chmlab <- chmlab()

  p <- ggplot(grodat_winter, aes(x = chemval, y = len)) +
    # geom_line(aes(group = station), colour = 'grey') +
    stat_smooth(method = 'lm', se =T, colour = 'red') +
    geom_label(aes(fill = mo, group = station, label = station), size = 3, colour = 'grey') +
    ylab("Average size") +
    xlab(chmlab) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      legend.title = element_blank(), 
      legend.position = 'top'
    )

  return(p)
  
}, height = 400, width = 400)

renderPrint({lapply(gromod_winter()['allchm'], summary)})
```

#### Summer cohort

```{r}
renderPlot({
  
  # inputs
  grodat_summer <- grodat_summer()
  chmlab <- chmlab()

  p <- ggplot(grodat_summer, aes(x = chemval, y = len)) +
    # geom_line(aes(group = station), colour = 'grey') +
    stat_smooth(method = 'lm', se =T, colour = 'red') +
    geom_label(aes(fill = mo, group = station, label = station), size = 3, colour = 'grey') +
    ylab("Average size") +
    xlab(chmlab) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      legend.title = element_blank(), 
      legend.position = 'top'
    )

  return(p)
  
}, height = 400, width = 400)

renderPrint({lapply(gromod_summer()['allchm'], summary)})
```

#### Response vs two variables

```{r}
column(12, 
       column(4, 
              selectInput('chm3', 'Select second chemistry variable:', choices = c('ara', 'araund', 'co3', 'ph', 'pco2', 'temp', 'oxy', 'chla'), selected = 'temp')
       ),
       
       column(4, 
              selectInput('val3', 'Select second chemistry value:', choices = c('min', 'ave', 'max', 'std', 'rng', 'dlt'))
       ),
       column(width = 4, 
              selectInput('modtyptmp2', "Include interaction:", choices = c(T, F))
       )
)
column(12, 
       column(4, 
              selectInput('flt2', 'Filter data by:', choices = c('nothing', 'Jul', 'Apr', 'Sep', '2014', '2015', '2016'))
              )
       )
```

```{r}
mod3dat <- reactive({  
  
  # inputs
  chm <- input$chm
  dis <- input$dis
  val <- input$val
  chm3 <- input$chm3
  val3 <- input$val3
  sta <- input$sta
  flt2 <- input$flt2
  
  validate(
    need(chm != chm3, 'Variable selections must be different')
  )
  
  dat <- chmdatsum %>% 
    filter(var %in% c(chm, chm3)) %>% 
    full_join(biodat, by = c('station', 'date', 'yr', 'cohortyr', 'mo')) %>% 
    select(-lon, -lat, -abu) %>% 
    gather('chemvar', 'chemval', ave:dlt) %>% 
    filter(var %in% chm & chemvar %in% val | var %in% chm3 & chemvar %in% val3) %>%
    filter(station %in% sta) %>% 
    select(-chemvar) %>% 
    spread(var, chemval)

  if(flt2 %in% c('Jul', 'Sep', 'Apr'))
    dat <- dat %>% filter(mo %in% flt2)

  if(flt2 %in% c('2014', '2015', '2016'))
    dat <- dat %>% filter(cohortyr %in% as.numeric(flt2))
  
  return(dat)
  
})

lmmod3 <- reactive({
  
  # input
  mod3dat <- mod3dat()
  modtyptmp2 <- input$modtyptmp2
  chm <- input$chm
  chm3 <- input$chm3
  
  frm <- paste('len ~', chm, '+', chm3)
  if(modtyptmp2)
    frm <- gsub('\\+', '*', frm)
  
  mod <- paste('lm(', frm, ', data = mod3dat)')
  mod <- eval(parse(text = mod))
  
  return(mod)
  
})

vifval2 <- reactive({
  
  # input
  lmmod3 <- lmmod3()
  chm3 <- input$chm3
  
  datin <- lmmod3$model[,-1] 
  vifval <- paste0('lm(', chm3, ' ~ ., data= datin)')
  vifval <- eval(parse(text = vifval)) %>% 
    summary %>% 
    .$r.squared
  vifval <- 1 / (1 - vifval)
  vifval <- round(vifval, 2) %>% 
    paste('VIF:', .)
  
  return(vifval)
  
})
 
# interaction plot
output$int3plt <- renderPlot({
  
  # input
  lmmod3 <- lmmod3()
  chmlab <- chmlab()
  chm3 <- input$chm3
  val3 <- input$val3
  
  # colors
  cols <- RColorBrewer::brewer.pal(9, 'RdBu')
  
  # get data to plot
  pldat <- get_pldat(lmmod3, chm3)
  pl1 <- pldat[[1]]
  pl2 <- pldat[[2]]
  
  # names for aes_string
  nms <- names(pl1)
  p1 <- ggplot() +
    geom_ribbon(data = pl1, aes_string(x = nms[3], ymin = 'lo', ymax = 'hi', group = nms[4]), alpha = 0.5, fill = 'grey') +
    geom_line(data = pl1, aes_string(x = nms[3], y = nms[1], group = nms[4], colour = nms[4]), size = 1) + 
    # geom_text(data = pl2, aes(x= x, y = y, label = lab), hjust = 0) +
    theme_bw() +
    scale_x_continuous(chmlab) +
    scale_colour_gradientn(paste(val3, chm3), colours = cols) + 
    scale_y_continuous('len')
  
  return(p1)
  
}, height = 300, width = 600) 

```

```{r}
plotOutput('int3plt')
```
<br><br><br>
```{r}
renderText({vifval2()})
renderPrint({summary(lmmod3())})
```

### Growth by dissolution {.tabset .tabset-pills}

#### Winter cohort

```{r}
renderPlot({
  
  # inputs
  grodat_winter <- grodat_winter()
  dis <- input$dis

  p <- ggplot(grodat_winter, aes(x = disval, y = len)) +
    # geom_line(aes(group = station), colour = 'grey') +
    stat_smooth(method = 'lm', se =T, colour = 'red') +
    geom_label(aes(fill = mo, group = station, label = station), size = 3, colour = 'grey') +
    ylab("Average size") +
    xlab(dis) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      legend.title = element_blank(), 
      legend.position = 'top'
    )

  return(p)
  
}, height = 400, width = 400)

renderPrint({lapply(gromod_winter()['alldis'], summary)})
```

#### Summer cohort

```{r}
renderPlot({
  
  # inputs
  grodat_summer <- grodat_summer()
  dis <- input$dis

  p <- ggplot(grodat_summer, aes(x = disval, y = len)) +
    # geom_line(aes(group = station), colour = 'grey') +
    stat_smooth(method = 'lm', se =T, colour = 'red') +
    geom_label(aes(fill = mo, group = station, label = station), size = 3, colour = 'grey') +
    ylab("Average size") +
    xlab(dis) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      legend.title = element_blank(), 
      legend.position = 'top'
    )

  return(p)
  
}, height = 400, width = 400)

renderPrint({lapply(gromod_summer()['alldis'], summary)})
```

### Abundance by environment {.tabset .tabset-pills}

```{r}
abudat <- reactive({

  # inputs
  chm <- input$chm
  val <- input$val
  sta <- input$sta

  dat <- chmdatsum %>% 
    filter(var %in% chm) %>% 
    full_join(biodat, by = c('station', 'date', 'yr', 'cohortyr', 'mo')) %>% 
    select(-lon, -lat, -var, -typ1, -typ2, -typ3, -len, -`% type II&III`) %>%
    gather('chemvar', 'chemval', ave:dlt) %>% 
    filter(chemvar %in% val) %>% 
    filter(station %in% sta)
  
  return(dat)
  
})

abumod <- reactive({
  
  # input
  abudat <- abudat()
  
  # complete model
  all <- lm(log10(abu) ~ chemval, data = abudat)
  
  # by cohort years
  bycohortyr <- abudat %>% 
    group_by(cohortyr) %>% 
    nest %>% 
    mutate(
      mod = purrr::map(data, function(x){ try({lm(log10(abu) ~ chemval, data = x)})})
    )
  bycohortyrout <- pull(bycohortyr, mod)
  names(bycohortyrout) <- pull(bycohortyr, cohortyr)
  
  # by seasons
  byseas <- abudat %>% 
    group_by(mo) %>% 
    nest %>% 
    mutate(
      mod = purrr::map(data, function(x){ try({lm(log10(abu) ~ chemval, data = x)})})
    )
  byseasout <- pull(byseas, mod)
  names(byseasout) <- pull(byseas, mo)

  # combine all    
  out <- c(
    all = list(all), 
    bycohortyrout,
    byseasout
  )
  
  return(out)
  
})
```

```{r}
renderPlot({
  
  # inputs
  abudat <- abudat()
  chmlab <- chmlab()

  p <- ggplot(abudat, aes(x = chemval, y = abu)) +
    # geom_line(aes(group = station), colour = 'grey') +
    stat_smooth(method = 'lm', se =T, colour = 'red') +
    geom_label(aes(fill = mo, group = station, label = station), size = 3, colour = 'grey') +
    ylab("Average size") +
    xlab(chmlab) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      legend.title = element_blank(), 
      legend.position = 'top'
    ) + 
    scale_y_log10('log10(abu)')

  return(p)
  
}, height = 400, width = 400)

renderPrint({lapply(abumod()['all'], summary)})

renderPlot({
  
  # inputs
  abudat <- abudat()
  chmlab <- chmlab()

  p <- ggplot(abudat, aes(x = chemval, y = abu)) +
    # geom_line(aes(group = station), colour = 'grey') +
    stat_smooth(method = 'lm', se = F, colour = 'red') +
    geom_label(aes(fill = mo, group = station, label = station), size = 3, colour = 'grey') +
    facet_grid(~ cohortyr) + #, scales = 'free') +
    ylab("Average size") +
    xlab(chmlab) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      legend.title = element_blank(), 
      legend.position = 'top'
    ) + 
    scale_y_log10('log10(abu)')

  return(p)
  
}, height = 300, width = 850)

renderPrint({lapply(abumod()[c('2014', '2015', '2016')], summary)})

renderPlot({
  
  # inputs
  abudat <- abudat()
  chmlab <- chmlab()

  p <- ggplot(abudat, aes(x = chemval, y = abu)) +
    # geom_line(aes(group = station), colour = 'grey') +
    stat_smooth(method = 'lm', se = F, colour = 'red') +
    geom_label(aes(fill = cohortyr, group = station, label = station), size = 3, colour = 'grey') +
    facet_grid(~ mo) + #, scales = 'free') +
    ylab("Average size") +
    xlab(chmlab) +
    theme_bw(base_family = 'serif', base_size = 16) +
    theme(
      strip.background = element_blank(), 
      legend.title = element_blank(), 
      legend.position = 'top'
    ) + 
    scale_y_log10('log10(abu)')

  return(p)
  
}, height = 300, width = 850)

renderPrint({lapply(abumod()[c('Jul', 'Sep', 'Apr')], summary)})
```

