---
title: "Pteropod Puget Sound time series"
author: ""
output: 
  html_document:
self_contained: yes
runtime: shiny
---

# Station similarities, between years

```{r, echo = F, warning = F, message = F}
knitr::opts_chunk$set(echo = F, warning = F, message = F)

library(tidyverse)
library(sf)
library(ggmap)
library(mapview)
library(lubridate)
library(readxl)
library(gridExtra)
library(vegan)
library(ggdendro)
library(dendextend)
library(mapview)
library(shiny)

load(file = 'data/biodat.RData')
load(file = 'data/chmdatsum.RData')
```

```{r, fig.height = 6, fig.width = 4}
locs <- chmdatsum %>% 
  select(station, lon, lat) %>% 
  unique 

# get the extent
dat_ext <- make_bbox(locs$lon, locs$lat, f = 0.05)

# get the base map using the extent
bsmap <-  get_map(location = dat_ext, maptype = 'toner-lite', source = 'stamen', zoom = 9)

# plot the basemap
ggmap(bsmap) + 
  geom_point(data = locs, aes(x = lon, y = lat), colour = 'tomato1', size = 8) +
  geom_text(data = locs, aes(x = lon, y = lat, label = station)) + 
  theme_bw(base_family = 'serif') +
  theme(axis.title = element_blank())
```

# {.tabset .tabset-pills}

Select either water quality or dissolution data for the plots and clustering analyses below.  

## Water quality 

### Monthly averages by water chemistry {.tabset}

The plots below show averages for salinity, temperature, omega, and oxygen across years for the same month.  Points are sized by magnitude of the average and colored with transparency for the standard deviation of the estimates.  
```{r}
toplo1a <- read_excel('raw/WOAC_data_5-1-2018_for_Nina.xlsx', sheet = 'ALL_DATA', na = c('', '-999')) %>% 
  select(Date_collected, STATION_NO, LATITUDE_DEC, LONGITUDE_DEC, NISKIN_NO, CTDTMP_DEG_C_ITS90, 
         CTDSAL_PSS78, CTDOXY_UMOL_KG_ADJ, `Omega Ar`) %>% 
  rename(
    date = Date_collected, 
    station = STATION_NO,
    lat = LATITUDE_DEC,
    lon = LONGITUDE_DEC,
    niskin = NISKIN_NO,
    temp = CTDTMP_DEG_C_ITS90,
    sal = CTDSAL_PSS78,
    ara = `Omega Ar`,
    oxy = CTDOXY_UMOL_KG_ADJ
  ) %>% 
  mutate(
    date = as.Date(date)
  ) %>% 
  gather('var', 'val', temp:ara) %>% 
  group_by(station) %>% 
  mutate(
    lat = mean(lat), 
    lon = mean(lon),
    yr = year(date), 
    mo = month(date)
  ) %>% 
  filter(mo %in% c(4, 7, 9, 10)) %>%
  mutate(
    mo = case_when(
      yr == 2014 & mo == 10 ~ 9,
      TRUE ~ mo
    ),
    mo = month(mo, label = T)
  )

# ggplot(toplo, aes(x = factor(station), y = val, fill = factor(yr))) + 
#   geom_boxplot(position = 'dodge') +
#   facet_grid(var ~ mo, scales = 'free_y')

# average across years by month
toplo1b <- toplo1a %>% 
  group_by(station, var, mo, lat, lon) %>% 
  summarise(
    valmn = mean(val, na.rm = T),
    valsd = sd(val, na.rm = T)
  ) %>% 
  ungroup %>% 
  mutate(
    mo = factor(mo, levels = c('Jul', 'Sep', 'Apr'))
  )

pbase <- theme_bw(base_size = 14, base_family = 'serif') + 
  theme(
    axis.title = element_blank(), 
    strip.background = element_blank() 
  ) 
```

#### Salinity

```{r, fig.height = 6, fig.width = 8}
salplo <- toplo1b %>% 
  filter(var %in% 'sal')
p1 <- ggmap(bsmap) +
  geom_point(data = salplo, aes(x = lon, y = lat, size = valmn, alpha = valsd), colour = 'tomato1') + 
  geom_point(data = salplo, aes(x = lon, y = lat, size = valmn), pch = 21) + 
  scale_size('Salinity (ave)', range = c(1, 10)) +
  scale_alpha('Salinity (sd)') +
  coord_map() + 
  facet_wrap(~mo) + 
  pbase +
  guides(alpha = guide_legend(override.aes = list(size=3)))

p1
```

#### Temperature

```{r, fig.height = 6, fig.width = 8}
tempplo <- toplo1b %>% 
  filter(var %in% 'temp')
p2 <- ggmap(bsmap) +
  geom_point(data = tempplo, aes(x = lon, y = lat, size = valmn, alpha = valsd), colour = 'tomato1') + 
  geom_point(data = tempplo, aes(x = lon, y = lat, size = valmn), pch = 21) + 
  scale_size('Temp (ave)', range = c(1, 10)) +
  scale_alpha('Temp (sd)') +
  coord_map() + 
  facet_wrap(~mo) + 
  pbase +
  guides(alpha = guide_legend(override.aes = list(size=3)))

p2
```

#### Omega

```{r, fig.height = 6, fig.width = 8}
araplo <- toplo1b %>% 
  filter(var %in% 'ara')
p3 <- ggmap(bsmap) +
  geom_point(data = araplo, aes(x = lon, y = lat, size = valmn, alpha = valsd), colour = 'tomato1') + 
  geom_point(data = araplo, aes(x = lon, y = lat, size = valmn), pch = 21) + 
  scale_size('Omega (ave)', range = c(1, 10)) +
  scale_alpha('Omega (sd)') +
  coord_map() + 
  facet_wrap(~mo) + 
  pbase +
  guides(alpha = guide_legend(override.aes = list(size=3)))

p3
```

#### Oxygen

```{r, fig.height = 6, fig.width = 8}
araplo <- toplo1b %>% 
  filter(var %in% 'oxy')
p3 <- ggmap(bsmap) +
  geom_point(data = araplo, aes(x = lon, y = lat, size = valmn, alpha = valsd), colour = 'tomato1') + 
  geom_point(data = araplo, aes(x = lon, y = lat, size = valmn), pch = 21) + 
  scale_size('Oxygen (ave)', range = c(1, 10)) +
  scale_alpha('Oxygen (sd)') +
  coord_map() + 
  facet_wrap(~mo) + 
  pbase +
  guides(alpha = guide_legend(override.aes = list(size=3)))

p3
```


### Clustering {.tabset}

Sites are clustered below based on water quality aggregations for the selected months.  

```{r}
selectInput('clmo1', 'Select month for clustering:', choices = c('Jul', 'Sep', 'Apr'), selected = 'Sep')
```

```{r}
toclst1 <- reactive({
  
  # input
  clmo1 <- input$clmo1

  out <- toplo1b %>%
    ungroup %>% 
    filter(mo %in% clmo1) %>% 
    select(station, var, valmn) %>% 
    spread(var, valmn) %>% 
    data.frame %>% 
    column_to_rownames('station') %>% 
    decostand(method = 'standardize')
  
  return(out)
  
})

# dissimilarity
dis1 <- reactive({
  
  # input
  toclst1 <- toclst1()
  
  out <- toclst1 %>% 
    vegdist(method = 'euclidean')
  
  return(out)
  
})
  
# clsuters
clst1 <- reactive({
  
  # input
  dis1 <- dis1()
  
  out <- dis1 %>%  
    hclust(method = 'average')
  
  return(out)
  
})

# cluster groups
clstgrps1 <- reactive({
  
  # input
  clst1 <- clst1()
  
  clstcut <- 3
  
  # get cut groups
  out <- cutree(clst1, k = clstcut)
  
  return(out)
  
})
```

#### Dendrogram

```{r}
renderPlot({

  # input
  clst1 <- clst1()
  clstgrps1 <- clstgrps1()

  clstcut <- max(clstgrps1)
  
  # get order
  clstord <- order.hclust(clst1) %>%
    clstgrps1[.] %>%
    unique
  
  # get colors, correct by order
  cols <- mapviewGetOption("vector.palette")(as.numeric(clstcut)) %>%
    .[clstord]
  
  p1 <- clst1 %>% 
    as.dendrogram %>% 
    set("branches_k_color", k = clstcut, value = cols) %>%
    set("labels_colors", k = clstcut, value = cols) %>%
    set("labels_cex", 0.8) 
  plot(p1)

})
```

#### Station dissimilarities

```{r}
renderPlot({

  # input
  dis1 <- dis1()
  clst1 <- clst1()
  clstgrps1 <- clstgrps1()

  # prep distance data to plot
  # long format of dist matrix
  toplo <- dis1 %>%
    as.matrix %>%
    as.data.frame %>%
    rownames_to_column('station') %>%
    gather('station2', 'dist', -station) %>%
    arrange(dist) %>%
    mutate(
      dist = ifelse(station == station2, NA, dist)
      )

  # get site order levels based on clustering
  sitfc <- clst1$labels[clst1$order]
  toplo <- toplo %>%
    mutate(
      station = factor(station, levels = sitfc),
      station2 = factor(station2, levels = sitfc)
    )

  # plot
  p <- ggplot(toplo) +
    geom_tile(aes(x = station, y = station2, fill = dist), colour = 'black') +
    scale_x_discrete('', expand = c(0, 0)) +
    scale_y_discrete('', expand = c(0, 0)) +
    scale_fill_gradient2('Dissimilarity between stations\nby water chemistry', low = 'lightblue', mid = 'white', high = 'tomato1', midpoint = 2) +
    guides(fill = guide_colourbar(barheight = 0.5, barwidth = 10, label.theme = element_text(size = 11, angle = 0))) +
    pbase

  # index values of cluster cuts
  brks <- clstgrps1 %>%
    .[clst1$order] %>%
    duplicated %>%
    `!` %>%
    which %>%
    `-` (0.5) %>%
    .[-1]

  p <- p +
    geom_vline(xintercept = brks, size = 1.5) +
    geom_hline(yintercept = brks, size = 1.5) +
    theme_bw(base_family = 'serif') +
    theme(
      legend.position = 'top',
      legend.direction = 'horizontal'
    )

  return(p)

}, height = 600, width = 700)
```

#### Map

```{r}
renderPlot({

  # input
  clst1 <- clst1()
  clstgrps1 <- clstgrps1()

  mapplo <- locs %>%
    mutate(clstgrps1 = clstgrps1)

  # plot the basemap
  ggmap(bsmap) +
    geom_point(data = mapplo, aes(x = lon, y = lat, fill = clstgrps1), pch = 21, size = 8, alpha = 0.7) +
    geom_text(data = mapplo, aes(x = lon, y = lat, label = station)) +
    scale_fill_viridis_c() +
    theme_bw(base_family = 'serif') +
    theme(
      axis.title = element_blank(),
      legend.position = 'none'
      )
  
}, height = 600, width = 400)
```

## Pteropod response

The plots below show averages for dissolution across years for the same month.  Points are sized by magnitude of the average and colored with transparency for the standard deviation of the estimates. 

### Monthly averages by response measures {.tabset}

```{r}
toplo2a <- biodat %>%
  select(cohortyr, mo, station, typ3) %>%
  left_join(locs, by = 'station') %>%
  gather('var', 'val', typ3) %>%
  rename(yr = cohortyr)

# ggplot(toplo, aes(x = factor(station), y = val, fill = factor(yr))) +
#   geom_boxplot(position = 'dodge') +
#   facet_grid(var ~ mo, scales = 'free_y')

# average across years by month
toplo2b <- toplo2a %>%
  group_by(station, var, mo, lat, lon) %>%
  summarise(
    valmn = mean(val, na.rm = T),
    valsd = sd(val, na.rm = T)
  ) %>%
  ungroup %>%
  mutate(
    mo = factor(mo, levels = c('Jul', 'Sep', 'Apr'))
  )

pbase <- theme_bw(base_size = 14, base_family = 'serif') +
  theme(
    axis.title = element_blank(),
    strip.background = element_blank()
  )
```

<!-- #### Type 1 -->

<!-- ```{r, fig.height = 6, fig.width = 8} -->
<!-- typ1plo <- toplo2b %>% -->
<!--   filter(var %in% 'typ1') -->
<!-- p1 <- ggmap(bsmap) + -->
<!--   geom_point(data = typ1plo, aes(x = lon, y = lat, size = valmn, alpha = valsd), colour = 'tomato1') + -->
<!--   geom_point(data = typ1plo, aes(x = lon, y = lat, size = valmn), pch = 21) + -->
<!--   scale_size('Type 1 dissolution (ave)', range = c(1, 10)) + -->
<!--   scale_alpha('Type 1 dissolution (sd)') + -->
<!--   coord_map() + -->
<!--   facet_wrap(~mo) + -->
<!--   pbase + -->
<!--   guides(alpha = guide_legend(override.aes = list(size=3))) -->

<!-- p1 -->
<!-- ``` -->

<!-- #### Type 2 -->

<!-- ```{r, fig.height = 6, fig.width = 8} -->
<!-- typ2plo <- toplo2b %>% -->
<!--   filter(var %in% 'typ2') -->
<!-- p2 <- ggmap(bsmap) + -->
<!--   geom_point(data = typ2plo, aes(x = lon, y = lat, size = valmn, alpha = valsd), colour = 'tomato1') + -->
<!--   geom_point(data = typ2plo, aes(x = lon, y = lat, size = valmn), pch = 21) + -->
<!--   scale_size('Type 2 dissolution (ave)', range = c(1, 10)) + -->
<!--   scale_alpha('Type 2 dissolution (sd)') + -->
<!--   coord_map() + -->
<!--   facet_wrap(~mo) + -->
<!--   pbase + -->
<!--   guides(alpha = guide_legend(override.aes = list(size=3))) -->

<!-- p2 -->
<!-- ``` -->

#### Type 3

```{r, fig.height = 6, fig.width = 8}
typ3plo <- toplo2b %>%
  filter(var %in% 'typ3')
p3 <- ggmap(bsmap) +
  geom_point(data = typ3plo, aes(x = lon, y = lat, size = valmn, alpha = valsd), colour = 'tomato1') +
  geom_point(data = typ3plo, aes(x = lon, y = lat, size = valmn), pch = 21) +
  scale_size('Type 3 dissolution (ave)', range = c(1, 10)) +
  scale_alpha('Type 3 dissolution (sd)') +
  coord_map() +
  facet_wrap(~mo) +
  pbase +
  guides(alpha = guide_legend(override.aes = list(size=3)))

p3
```

### Clustering {.tabset}

Sites are clustered below based on dissolutuion aggregations for the selected months.  

```{r}
selectInput('clmo2', 'Select month for clustering:', choices = c('Jul', 'Sep', 'Apr'), selected = 'Sep')
```

```{r}
toclst2 <- reactive({

  # input
  clmo2 <- input$clmo2

  out <- toplo2b %>%
    ungroup %>%
    filter(mo %in% clmo2) %>%
    select(station, var, valmn) %>%
    spread(var, valmn) %>%
    data.frame %>%
    column_to_rownames('station')

  return(out)

})

# dissimilarity
dis2 <- reactive({

  # input
  toclst2 <- toclst2()

  out <- toclst2 %>%
    vegdist(method = 'euclidean')

  return(out)

})

# clsuters
clst2 <- reactive({

  # input
  dis2 <- dis2()

  out <- dis2 %>%
    hclust(method = 'average')

  return(out)

})

# cluster groups
clstgrps2 <- reactive({

  # input
  clst2 <- clst2()

  clstcut <- 3

  # get cut groups
  out <- cutree(clst2, k = clstcut)

  return(out)

})
```

#### Dendrogram

```{r}
renderPlot({

  # input
  clst2 <- clst2()
  clstgrps2 <- clstgrps2()

  clstcut <- max(clstgrps2)

  # get order
  clstord <- order.hclust(clst2) %>%
    clstgrps2[.] %>%
    unique

  # get colors, correct by order
  cols <- mapviewGetOption("vector.palette")(as.numeric(clstcut)) %>%
    .[clstord]

  p1 <- clst2 %>%
    as.dendrogram %>%
    set("branches_k_color", k = clstcut, value = cols) %>%
    set("labels_colors", k = clstcut, value = cols) %>%
    set("labels_cex", 0.8)
  plot(p1)

})
```

#### Station dissimilarities

```{r}
renderPlot({

  # input
  dis2 <- dis2()
  clst2 <- clst2()
  clstgrps2 <- clstgrps2()

  # prep distance data to plot
  # long format of dist matrix
  toplo <- dis2 %>%
    as.matrix %>%
    as.data.frame %>%
    rownames_to_column('station') %>%
    gather('station2', 'dist', -station) %>%
    arrange(dist) %>%
    mutate(
      dist = ifelse(station == station2, NA, dist)
      )

  # get site order levels based on clustering
  sitfc <- clst2$labels[clst2$order]
  toplo <- toplo %>%
    mutate(
      station = factor(station, levels = sitfc),
      station2 = factor(station2, levels = sitfc)
    )

  # plot
  p <- ggplot(toplo) +
    geom_tile(aes(x = station, y = station2, fill = dist), colour = 'black') +
    scale_x_discrete('', expand = c(0, 0)) +
    scale_y_discrete('', expand = c(0, 0)) +
    scale_fill_gradient2('Dissimilarity between stations\nby dissolution', low = 'lightblue', mid = 'white', high = 'tomato1', midpoint = 2) +
    guides(fill = guide_colourbar(barheight = 0.5, barwidth = 10, label.theme = element_text(size = 11, angle = 0))) +
    pbase

  # index values of cluster cuts
  brks <- clstgrps2 %>%
    .[clst2$order] %>%
    duplicated %>%
    `!` %>%
    which %>%
    `-` (0.5) %>%
    .[-1]

  p <- p +
    geom_vline(xintercept = brks, size = 1.5) +
    geom_hline(yintercept = brks, size = 1.5) +
    theme_bw(base_family = 'serif') +
    theme(
      legend.position = 'top',
      legend.direction = 'horizontal'
    )

  return(p)

}, height = 600, width = 700)
```

#### Map

```{r}
renderPlot({

  # input
  clst2 <- clst2()
  clstgrps2 <- clstgrps2()

  mapplo <- locs %>%
    mutate(clstgrps2 = clstgrps2)

  # plot the basemap
  ggmap(bsmap) +
    geom_point(data = mapplo, aes(x = lon, y = lat, fill = clstgrps2), pch = 21, size = 8, alpha = 0.7) +
    geom_text(data = mapplo, aes(x = lon, y = lat, label = station)) +
    scale_fill_viridis_c() +
    theme_bw(base_family = 'serif') +
    theme(
      axis.title = element_blank(),
      legend.position = 'none'
      )

}, height = 600, width = 400)
```